---
title: "Diseases Mortality changes due to the effects of Medical Reforms"
output: html_notebook
---
Downloading Libraries.
```{r}
library(ggplot2)
library(pracma)
library(pastecs)
library(MASS)
```

Collecting data and forming dataframes.
```{r}
data.mortality_germany <- read.csv("DATASET_GERMANY/DeathRateGermany - Sheet1.csv")
data.life_expectancy_germany <- read.csv("DATASET_GERMANY/LifeExpectancyGermany - Sheet1.csv")

data.mortality_norway <- read.csv("DATASET_NORWAY/DeathRateNorway - Sheet1.csv")
data.life_expectancy_norway <- read.csv("DATASET_NORWAY/LifeExpectancyNorway - Sheet1.csv")

data.mortality_germany_r <- data.frame("Year" = rev(data.mortality_germany$Year), "Death.Rate" = rev(data.mortality_germany$Death.Rate), "Growth.Rate" = rev(data.mortality_germany$Growth.Rate))
data.life_expectancy_germany_r <- data.frame("Year" = rev(data.life_expectancy_germany$Year), "Life.Expectancy" = rev(data.life_expectancy_germany$Life.Expectancy), "Growth.Rate" = rev(data.life_expectancy_germany$Growth.rate))

data.mortality_norway_r <- data.frame("Year" = rev(data.mortality_norway$Year), "Death.Rate" = rev(data.mortality_norway$Death.Rate), "Growth.Rate" = rev(data.mortality_norway$Growth.Rate))
data.life_expectancy_norway_r <- data.frame("Year" = rev(data.life_expectancy_norway$Year), "Life.Expectancy" = rev(data.life_expectancy_norway$Life.Expectancy), "Growth.Rate" = rev(data.life_expectancy_norway$Growth.Rate))

data.mortality_norway_r
data.life_expectancy_norway_r
```

Running linear regression to see if the data has a trend.
```{r}
year_m <- data.mortality_germany_r$Year
death_rate_germany <- data.mortality_germany_r$Death.Rate
death_rate_norway <- data.mortality_norway_r$Death.Rate
summary(lm(death_rate_germany~year_m))
summary(lm(death_rate_norway~year_m))
```

By the p-values for variable year_m, we observe, that the Germany's mortality rate has to be detrended. The plot below shows the results. 
```{r}
data.germany_mort_dentrended <- merge(data.mortality_germany_r[, c("Year", "Death.Rate")], data.mortality_germany_r[, c("Year", "Death.Rate")], by.x="Year", by.y="Year")

ggplot(data.germany_mort_dentrended, aes(x=Year)) + geom_line(aes(y=Death.Rate.x, color="With Trend")) + geom_line(aes(y=detrend(Death.Rate.y), color="Detrended")) + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("", values=c("With Trend"="steelblue", "Detrended"="firebrick4")) + ggtitle("Germany Total Mortality Rate")
```

Forming diseases mortality dataframes.
```{r}
data.diseases_mortality_germany <- read.csv("DATASET_GERMANY/MortalityDiseasesGermany - Sheet1.csv")

data.diseases_mortality_germany_r <- data.frame(
  "Year" = rev(data.diseases_mortality_germany$Year), "Icheamic.Heart.Disease" = rev(data.diseases_mortality_germany$Icheamic.Heart.Disease), "Mental.and.behavioural.disorders" = rev(data.diseases_mortality_germany$Mental.and.behavioural.disorders), "Diseases.of.the.nervous.system" = rev(data.diseases_mortality_germany$Diseases.of.the.nervous.system), "Malignant.neoplasm.of.breast..female" = rev(data.diseases_mortality_germany$Malignant.neoplasm.of.breast..female), "Diabetes.mellitus" = rev(data.diseases_mortality_germany$Diabetes.mellitus), "Leukaemia" = rev(data.diseases_mortality_germany$Leukaemia), "Alzheimer" = rev(data.diseases_mortality_germany$Alzheimer), "Pneumonia" = rev(data.diseases_mortality_germany$Pneumonia), "TotalDeaths"=rev(data.diseases_mortality_germany$Total.Deaths))

data.diseases_mortality_germany_r
```

```{r}
data.diseases_mortality_norway <- read.csv("DATASET_NORWAY/MortalityDiseasesNorway - Sheet1.csv")[1:26,]

data.diseases_mortality_norway_r <- data.frame(
  "Year" = rev(data.diseases_mortality_norway$Year), "Icheamic.Heart.Disease" = rev(data.diseases_mortality_norway$Icheamic.Heart.Disease), "Mental.and.behavioural.disorders" = rev(data.diseases_mortality_norway$Mental.and.behavioural.disorders), "Diseases.of.the.nervous.system" = rev(data.diseases_mortality_norway$Diseases.of.the.nervous.system), "Malignant.neoplasm.of.breast..female" = rev(data.diseases_mortality_norway$Malignant.neoplasm.of.breast..female), "Diabetes.mellitus" = rev(data.diseases_mortality_norway$Diabetes.mellitus), "Leukaemia" = rev(data.diseases_mortality_norway$Leukaemia), "Alzheimer" = rev(data.diseases_mortality_norway$Alzheimer), "Pneumonia" = rev(data.diseases_mortality_norway$Pneumonia), "TotalDeaths"=rev(data.diseases_mortality_norway$Total.Deaths))

data.diseases_mortality_norway_r
```

```{r}
data.diseases_mortality_ukraine <- read.csv("DATASET_UKRAINE/MortalityDiseasesUkraine - Sheet1.csv")

data.diseases_mortality_ukraine_r <- data.frame("Year" = rev(data.diseases_mortality_ukraine$Year), "Icheamic.Heart.Disease" = rev(data.diseases_mortality_ukraine$Icheamic.Heart.Disease), "Mental.and.behavioural.disorders" = rev(data.diseases_mortality_ukraine$Mental.and.behavioural.disorders), "Diseases.of.the.nervous.system" = rev(data.diseases_mortality_ukraine$Diseases.of.the.nervous.system), "Malignant.neoplasm.of.breast..female" = rev(data.diseases_mortality_ukraine$Malignant.neoplasm.of.breast..female), "Diabetes.mellitus" = rev(data.diseases_mortality_ukraine$Diabetes.mellitus), "Leukaemia" = rev(data.diseases_mortality_ukraine$Leukaemia), "Pneumonia" = rev(data.diseases_mortality_ukraine$Pneumonia), "TotalDeaths"=rev(data.diseases_mortality_ukraine$Total.Deaths))

data.diseases_mortality_ukraine_r
```

```{r}
data.diseases_mortality_belarus <- read.csv("DATASET_BELARUS/MortalityDiseasesBelarus - Sheet1.csv")

data.diseases_mortality_belarus_r <- data.frame("Year" = rev(data.diseases_mortality_belarus$Year), "Icheamic.Heart.Disease" = rev(data.diseases_mortality_belarus$Icheamic.Heart.Disease), "Mental.and.behavioural.disorders" = rev(data.diseases_mortality_belarus$Mental.and.behavioural.disorders), "Diseases.of.the.nervous.system" = rev(data.diseases_mortality_belarus$Diseases.of.the.nervous.system), "Malignant.neoplasm.of.breast..female" = rev(data.diseases_mortality_belarus$Malignant.neoplasm.of.breast..female), "Diabetes.mellitus" = rev(data.diseases_mortality_belarus$Diabetes.mellitus), "Leukaemia" = rev(data.diseases_mortality_belarus$Leukaemia), "Pneumonia" = rev(data.diseases_mortality_belarus$Pneumonia), "TotalDeaths"=rev(data.diseases_mortality_belarus$Total.Deaths))

data.diseases_mortality_belarus_r
```

Assigning values to the variables.
```{r}
year <- data.diseases_mortality_germany_r$Year
icheamic_heart_disease_mortality_germany <- data.diseases_mortality_germany_r$Icheamic.Heart.Disease
mental_behavioral_disorders_mortality_germany <- data.diseases_mortality_germany_r$Mental.and.behavioural.disorders
nervous_system_mortality_germany <- data.diseases_mortality_germany_r$Diseases.of.the.nervous.system
breast_cancer_mortality_germany <- data.diseases_mortality_germany_r$Malignant.neoplasm.of.breast..female
diabetes_mortality_germany <- data.diseases_mortality_germany_r$Diabetes.mellitus
leukaemia_mortality_germany <- data.diseases_mortality_germany_r$Leukaemia
alzheimer_mortality_germany <- data.diseases_mortality_germany_r$Alzheimer
pneumonia_mortality_germany <- data.diseases_mortality_germany_r$Pneumonia
total_mortality_germany <- data.diseases_mortality_germany_r$TotalDeaths

diseases_mortality_set <- c(icheamic_heart_disease_mortality_germany, mental_behavioral_disorders_mortality_germany, nervous_system_mortality_germany, breast_cancer_mortality_germany, diabetes_mortality_germany, leukaemia_mortality_germany, alzheimer_mortality_germany, pneumonia_mortality_germany)

column_names <- colnames(data.diseases_mortality_germany_r)[2:10]
```


```{r}
icheamic_heart_disease_mortality_norway <- data.diseases_mortality_norway_r$Icheamic.Heart.Disease
mental_behavioral_disorders_mortality_norway <- data.diseases_mortality_norway_r$Mental.and.behavioural.disorders
nervous_system_mortality_norway <- data.diseases_mortality_norway_r$Diseases.of.the.nervous.system
breast_cancer_mortality_norway <- data.diseases_mortality_norway_r$Malignant.neoplasm.of.breast..female
diabetes_mortality_norway <- data.diseases_mortality_norway_r$Diabetes.mellitus
leukaemia_mortality_norway <- data.diseases_mortality_norway_r$Leukaemia
alzheimer_mortality_norway <- data.diseases_mortality_norway_r$Alzheimer
pneumonia_mortality_norway <- data.diseases_mortality_norway_r$Pneumonia
total_mortality_norway <- data.diseases_mortality_norway_r$TotalDeaths
```

After plotting the values of icheamic heart disease mortality we can logically assume, that the data follows a downward trend, so we are going to check this by running linear regression.
```{r}
plot(icheamic_heart_disease_mortality_germany)
trend_check <- lm(icheamic_heart_disease_mortality_germany ~ year)
summary(trend_check)
predicted <- predict(trend_check)
abline(trend_check)
```

By the results we observe the high significance of time. Now, we are going to use built-in trend test to see whether all of the diseases data follow some trend.
```{r}
for (i in c(1:9)) {
  cat(column_names[i], "\n")
  print(trend.test(data.diseases_mortality_germany_r[,column_names[i]]))
  cat("\n\n")
}
```

Cannot compute exact p-value with ties warning. Ties - if dataset has equal values. (https://r.789695.n4.nabble.com/What-are-ties-Wilcox-u-test-td857059.html)
```{r}
for (i in c(1:8)) {
  cat(column_names[i], "\n")
  print(trend.test(data.diseases_mortality_norway_r[,column_names[i]]))
  cat("\n\n")
}
```

For Norway - Diabetes time series does not need a detrending because of a large p-value. The other ones require the detrending.

DETRENDING
```{r}
icheamic_heart_disease_mortality_germany_detrended <- detrend(icheamic_heart_disease_mortality_germany)
mental_behavioral_disorders_mortality_germany_detrended <- detrend(mental_behavioral_disorders_mortality_germany)
nervous_system_mortality_germany_detrended <- detrend(nervous_system_mortality_germany)
breast_cancer_mortality_germany_detrended <- detrend(breast_cancer_mortality_germany)
diabetes_mortality_germany_detrended <- detrend(diabetes_mortality_germany)
leukaemia_mortality_germany_detrended <- detrend(leukaemia_mortality_germany)
alzheimer_mortality_germany_detrended <- detrend(alzheimer_mortality_germany)
pneumonia_mortality_germany_detrended <- detrend(pneumonia_mortality_germany)
total_mortality_germany_detrended <- detrend(total_mortality_germany)
```

```{r}
icheamic_heart_disease_mortality_norway_detrended <- detrend(icheamic_heart_disease_mortality_norway)
mental_behavioral_disorders_mortality_norway_detrended <- detrend(mental_behavioral_disorders_mortality_norway)
nervous_system_mortality_norway_detrended <- detrend(nervous_system_mortality_norway)
breast_cancer_mortality_norway_detrended <- detrend(breast_cancer_mortality_norway)
diabetes_mortality_norway_detrended <- detrend(diabetes_mortality_norway)
leukaemia_mortality_norway_detrended <- detrend(leukaemia_mortality_norway)
alzheimer_mortality_norway_detrended <- detrend(alzheimer_mortality_norway)
pneumonia_mortality_norway_detrended <- detrend(pneumonia_mortality_norway)
total_mortality_norway_detrended <- detrend(total_mortality_norway)
```

LINEAR REGRESSION ON THE DEATH RATE BEING INFLUENCED BY THE DISEASES MORTALITY 
```{r}
death_rate_germany_1990_2015 <- death_rate_germany[41:66]
overall_mortality_model_germany <- lm(death_rate_germany_1990_2015~icheamic_heart_disease_mortality_germany+nervous_system_mortality_germany+diabetes_mortality_germany+nervous_system_mortality_germany+leukaemia_mortality_germany+alzheimer_mortality_germany+pneumonia_mortality_germany+breast_cancer_mortality_germany)
summary(overall_mortality_model_germany)
```
THE SAME REGRESSION BUT DETRENDED
```{r}
death_rate_germany_1990_2015_detrended <- detrend(death_rate_germany_1990_2015)
overall_mortality_model_germany_detrended <- lm(death_rate_germany_1990_2015_detrended~icheamic_heart_disease_mortality_germany_detrended+nervous_system_mortality_germany_detrended+diabetes_mortality_germany_detrended+nervous_system_mortality_germany_detrended+leukaemia_mortality_germany_detrended+alzheimer_mortality_germany_detrended+pneumonia_mortality_germany_detrended+breast_cancer_mortality_germany_detrended)
summary(overall_mortality_model_germany_detrended)
```

By results of the detrended regression we can see, that the icheamic heart disease, alzheimer, pneumonia, breast cancer mortality rates in our case influence the total death rate the most. 
```{r}
life_rate_germany <- data.life_expectancy_germany_r$Life.Expectancy
life_rate_norway <- data.life_expectancy_norway_r$Life.Expectancy
summary(lm(life_rate_germany~year_m))
summary(lm(life_rate_norway~year_m))
```

The results show that the life expectancy rate has a time trend both in Germany and in Norway and thus has to be detrended.

```{r}
data.life_dentrended <- merge(data.life_expectancy_germany_r[, c("Year", "Life.Expectancy")], data.life_expectancy_norway_r[, c("Year", "Life.Expectancy")], by.x="Year", by.y="Year")

ggplot(data.life_dentrended, aes(x=Year)) + geom_line(aes(y=Life.Expectancy.x, color="Germany")) + geom_line(aes(y=Life.Expectancy.y, color="Norway")) + geom_line(aes(y=detrend(Life.Expectancy.x), color="Germany"), linetype = "twodash") + geom_line(aes(y=detrend(Life.Expectancy.y), color="Norway"), linetype = "twodash") + scale_y_continuous(name="Life Expectancy") + scale_color_manual("Country", values=c("Germany"="steelblue", "Norway"="firebrick4")) + ggtitle("Life Expectancy Rate")
```

Run regression, where independent variables are mortalities from particular diseases and dependent is the life expectancy.
```{r}
life_rate_germany_1990_2015 <- life_rate_germany[41:66]
life_expectancy_model_germany <- lm(life_rate_germany_1990_2015~icheamic_heart_disease_mortality_germany+nervous_system_mortality_germany+diabetes_mortality_germany+nervous_system_mortality_germany+leukaemia_mortality_germany+alzheimer_mortality_germany+pneumonia_mortality_germany+breast_cancer_mortality_germany)
summary(life_expectancy_model_germany)
```
```{r}
life_rate_germany_1990_2015_detrended <- detrend(life_rate_germany_1990_2015) 
life_expectancy_model_germany_detrended <- lm(life_rate_germany_1990_2015_detrended~icheamic_heart_disease_mortality_germany_detrended+nervous_system_mortality_germany_detrended+diabetes_mortality_germany_detrended+nervous_system_mortality_germany_detrended+leukaemia_mortality_germany_detrended+alzheimer_mortality_germany_detrended+pneumonia_mortality_germany_detrended+breast_cancer_mortality_germany_detrended)
summary(life_expectancy_model_germany_detrended)
```

The results showed that it is slightly influenced by pneumonia mortality and Alzheimer disease. 

https://academic.oup.com/heapol/article/27/1/76/659122

SEGMENTED LINEAR REGRESSION
To see the effect of the reform
```{r}
time <- c(1:length(year)) #as we have data from 1990 to 2015
reform <- c(rep(0, 13), rep(1, 13)) # 0 - year with no reform, 1 - year with reform
postslope <- c(rep(0, 13), c(1:13))

#WITH TREND
icheamic_segmented_model <- lm(icheamic_heart_disease_mortality_germany~time+reform+postslope)
summary(icheamic_segmented_model)
```

```{r}
breast_cancer_mortality_germany_model <- lm(breast_cancer_mortality_germany~time+reform+postslope)
summary(breast_cancer_mortality_germany_model)
```

```{r}
diabetes_mortality_germany_model <- lm(diabetes_mortality_germany~time+reform+postslope)
summary(diabetes_mortality_germany_model)
```

FORMING A DATASET TO PLOT GERMANY AND NORWAY DATA 
```{r}
norway_germany <- merge(data.diseases_mortality_germany_r, data.diseases_mortality_norway_r, by.x="Year", by.y="Year")
norway_germany
```

Parallel Trend Assumption
It requires that in the absence of treatment, the difference between the 'treatment' and 'control' group is constant over time. Although there is no statistical test for this assumption, visual inspection is useful when you have observations over many time points.

VISUAL INSPECTION OF PARALLEL TREND TO RUN DIF IN DIF]
Plot predicted trend line of the diseases. 
```{r}
ggplot(norway_germany, aes(x=year)) + geom_line(aes(y=icheamic_heart_disease_mortality_germany, color="Germany")) + geom_line(aes(y=icheamic_heart_disease_mortality_norway, color="Norway")) + geom_line(aes(y=icheamic_heart_disease_mortality_germany_detrended, color="Germany"), linetype="twodash") + geom_line(aes(y=icheamic_heart_disease_mortality_norway_detrended, color="Norway"), linetype="twodash") + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkred", "Norway"="steelblue")) + ggtitle("Icheamic Heart Disease") + geom_vline(xintercept = 2002, color="red")

ggplot(norway_germany[6:13,], aes(x=year[6:13])) + geom_line(aes(y=predict(lm(icheamic_heart_disease_mortality_germany[6:13]~c(1995:2002))), color="Germany")) + geom_line(aes(y=predict(lm(icheamic_heart_disease_mortality_norway[6:13]~c(1995:2002))), color="Norway")) + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkred", "Norway"="steelblue")) + ggtitle("Icheamic Heart Disease")
```

```{r}
ggplot(norway_germany, aes(x=year)) + geom_line(aes(y=diabetes_mortality_germany, color="Germany")) + geom_line(aes(y=diabetes_mortality_norway, color="Norway")) + geom_line(aes(y=diabetes_mortality_germany_detrended, color="Germany"), linetype="twodash") + geom_line(aes(y=diabetes_mortality_norway_detrended, color="Norway"), linetype="twodash") + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkslateblue", "Norway"="firebrick4")) + ggtitle("Diabetes") + geom_vline(xintercept = 2002, color="red")

ggplot(norway_germany[1:13,], aes(x=year[1:13])) + geom_line(aes(y=predict(lm(diabetes_mortality_germany[1:13]~c(1990:2002))), color="Germany")) + geom_line(aes(y=predict(lm(diabetes_mortality_norway[1:13]~c(1990:2002))), color="Norway")) + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkred", "Norway"="steelblue")) + ggtitle("Diabetes")
```

```{r}
ggplot(norway_germany, aes(x=year)) + geom_line(aes(y=breast_cancer_mortality_germany, color="Germany")) + geom_line(aes(y=breast_cancer_mortality_norway, color="Norway")) + geom_line(aes(y=breast_cancer_mortality_germany_detrended, color="Germany"), linetype="twodash") + geom_line(aes(y=breast_cancer_mortality_norway_detrended, color="Norway"), linetype="twodash") + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="indianred4", "Norway"="orchid3")) + ggtitle("Breast Cancer") + geom_vline(xintercept = 2002, color="red")

ggplot(norway_germany[1:13,], aes(x=year[1:13])) + geom_line(aes(y=predict(lm(breast_cancer_mortality_germany[1:13]~c(1990:2002))), color="Germany")) + geom_line(aes(y=predict(lm(breast_cancer_mortality_norway[1:13]~c(1990:2002))), color="Norway")) + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkred", "Norway"="steelblue")) + ggtitle("Breast Cancer")
```

VISUAL OVERVIEW OF THE DISEASES NOT INFLUENCED BY THE REFORM
```{r}
ggplot(norway_germany, aes(x=year)) + geom_line(aes(y=mental_behavioral_disorders_mortality_germany, color="Germany")) + geom_line(aes(y=mental_behavioral_disorders_mortality_norway, color="Norway")) + geom_line(aes(y=mental_behavioral_disorders_mortality_germany_detrended, color="Germany"), linetype="twodash") + geom_line(aes(y=mental_behavioral_disorders_mortality_norway_detrended, color="Norway"), linetype="twodash") + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkred", "Norway"="steelblue")) + ggtitle("Mental Illnesses") + geom_vline(xintercept = 2002, color="red")

ggplot(norway_germany, aes(x=year)) + geom_line(aes(y=nervous_system_mortality_germany, color="Germany")) + geom_line(aes(y=nervous_system_mortality_norway, color="Norway")) + geom_line(aes(y=nervous_system_mortality_germany_detrended, color="Germany"), linetype="twodash") + geom_line(aes(y=nervous_system_mortality_norway_detrended, color="Norway"), linetype="twodash") + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkred", "Norway"="steelblue")) + ggtitle("Nervous System Diseases") + geom_vline(xintercept = 2002, color="red")

ggplot(norway_germany, aes(x=year)) + geom_line(aes(y=leukaemia_mortality_germany, color="Germany")) + geom_line(aes(y=leukaemia_mortality_norway, color="Norway")) + geom_line(aes(y=leukaemia_mortality_germany_detrended, color="Germany"), linetype="twodash") + geom_line(aes(y=leukaemia_mortality_norway_detrended, color="Norway"), linetype="twodash") + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkred", "Norway"="steelblue")) + ggtitle("Leukaemia") + geom_vline(xintercept = 2002, color="red")

ggplot(norway_germany, aes(x=year)) + geom_line(aes(y=alzheimer_mortality_germany, color="Germany")) + geom_line(aes(y=alzheimer_mortality_norway, color="Norway")) + geom_line(aes(y=alzheimer_mortality_germany_detrended, color="Germany"), linetype="twodash") + geom_line(aes(y=alzheimer_mortality_norway_detrended, color="Norway"), linetype="twodash") + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkred", "Norway"="steelblue")) + ggtitle("Alzheimer") + geom_vline(xintercept = 2002, color="red")

ggplot(norway_germany, aes(x=year)) + geom_line(aes(y=pneumonia_mortality_germany, color="Germany")) + geom_line(aes(y=pneumonia_mortality_norway, color="Norway")) + geom_line(aes(y=pneumonia_mortality_germany_detrended, color="Germany"), linetype="twodash") + geom_line(aes(y=pneumonia_mortality_norway_detrended, color="Norway"), linetype="twodash") + scale_y_continuous(name="Mortality per 100.000") + scale_color_manual("Country", values=c("Germany"="darkred", "Norway"="steelblue")) + ggtitle("Pneumonia") + geom_vline(xintercept = 2002, color="red")
```


DIF IN DIF  (https://www.mailman.columbia.edu/research/population-health-methods/difference-difference-estimation)
```{r}
did.dataframe <- data.frame("Country"=c(rep("Germany",26), rep("Norway", 26)), "Treatment"=c(rep(1, 26), rep(0, 26)), "Year"=c(c(1990:2015), c(1990:2015)), "YearDummy" = c(rep(0, 13), rep(1, 13), rep(0, 13), rep(1, 13)), "MortalityIcheamic"=c(data.diseases_mortality_germany_r$Icheamic.Heart.Disease, data.diseases_mortality_norway_r$Icheamic.Heart.Disease), "MortalityDiabetes"=c(data.diseases_mortality_germany_r$Diabetes.mellitus, data.diseases_mortality_norway_r$Diabetes.mellitus), "MortalityBreastCancer"=c(data.diseases_mortality_germany_r$Malignant.neoplasm.of.breast..female, data.diseases_mortality_norway_r$Malignant.neoplasm.of.breast..female))

did.dataframe
```

```{r}
did.dataframe_icheamic <- data.frame("Country"=c(rep("Germany",21), rep("Norway", 21)), "Treatment"=c(rep(1, 21), rep(0, 21)), "Year"=c(c(1995:2015), c(1995:2015)), "YearDummy" = c(rep(0, 8), rep(1, 13), rep(0, 8), rep(1, 13)), "MortalityIcheamic"=c(data.diseases_mortality_germany_r$Icheamic.Heart.Disease[6:26], data.diseases_mortality_norway_r$Icheamic.Heart.Disease[6:26]))
did.dataframe_icheamic
```


```{r}
did.mortality_icheamic <- did.dataframe_icheamic$MortalityIcheamic
did.treatment_icheamic <- did.dataframe_icheamic$Treatment
did.year_dummy_icheamic <- did.dataframe_icheamic$YearDummy
did.d_icheamic <- did.treatment_icheamic*did.year_dummy_icheamic
did.model_icheamic_lm <- lm(did.mortality_icheamic~did.treatment_icheamic+did.year_dummy_icheamic+did.d_icheamic)
did.model_icheamic_rlm <- rlm(did.mortality_icheamic~did.treatment_icheamic+did.year_dummy_icheamic+did.d_icheamic)
summary(did.model_icheamic_lm)
summary(did.model_icheamic_rlm)
```
t-value > critical value
1.01 < 1.68

```{r}
did.mortality_diabetes <- did.dataframe$MortalityDiabetes
did.treatment <- did.dataframe$Treatment
did.year_dummy <- did.dataframe$YearDummy
did.d <- did.treatment*did.year_dummy
did.model_diabetes_lm <- lm(did.mortality_diabetes~did.treatment+did.year_dummy+did.d)
did.model_diabetes_rlm <- rlm(did.mortality_diabetes~did.treatment+did.year_dummy+did.d)
summary(did.model_diabetes_lm)
summary(did.model_diabetes_rlm)
```

```{r}
did.mortality_breast_cancer <- did.dataframe$MortalityBreastCancer
did.model_breast_cancer_lm <- lm(did.mortality_breast_cancer~did.treatment+did.year_dummy+did.d)
did.model_breast_cancer_rlm <- rlm(did.mortality_breast_cancer~did.treatment+did.year_dummy+did.d)
summary(did.model_breast_cancer_lm)
summary(did.model_breast_cancer_rlm)
```

```{r}
data.diseases_mortality_belarus_r[complete.cases(data.diseases_mortality_belarus_r),]
data.diseases_mortality_ukraine_r[complete.cases(data.diseases_mortality_ukraine_r),]
```



















