---
title: "Investigation on which factors influence whether a country has a social government-covered medical insurance "
output: html_notebook
---
In this paper we want to investigate whether Ukraine is ready for introduction of mandatory social insurance. Based on experience of other countries we will perform a binary model that will depend on variety of macroeconomic factors as well as the situation with health in the country. 

Data description:
- General government deficit is measured as % of GDP : "Net lending" means that government has a surplus, and is providing financial resources to other sectors, while "net borrowing" means that government has a deficit, and requires financial resources from other sectors. This indicator is measured as a percentage of GDP.
- Household spending is the amount of final consumption expenditure made by resident households to meet their everyday needs, such as food, clothing, housing (rent), energy, etc. It is measured in Million US dollars. (can be measured as  % of GDP)
- Gross national income (GNI) is defined as gross domestic product, plus net receipts from abroad of compensation of employees, property income and net taxes less subsidies on production. This indicator is measures in Million US dollars. 
- Gross domestic product (GDP) is the standard measure of the value added created through the production of goods and services in a country during a certain period, measured in Million US dollars.
- Inflation measured by consumer price index (CPI) is defined as the change in the prices of a basket of goods and services that are typically purchased by specific groups of households. Inflation is measured in terms of the annual growth rate (percentage).
- Unemployment rate is the number of unemployed people as a percentage of the labour force.

List of countries who at 2016 had government-issued insurance:
Australia
Canada
Czech Republic
Denmark
Finland
Greece
Hungary
Ireland
Italy
Japan
Norway 
UK
Austria
Slovakia
Poland
Turkey

Full : https://en.wikipedia.org/wiki/List_of_countries_by_health_insurance_coverage 

Loading the data about those who had as well as those who did not.
```{r}
gdp <- read.csv("GDP.csv")
gni = read.csv("GNI.csv")
inflation = read.csv("INFLATION.csv")
spending = read.csv("HOUSEHOLD SPENDING.csv")
deficit = read.csv("DEFICIT.csv")
unemloyment = read.csv("UNEMPLOYMENT.csv")
gdp
```
Separating those who have and forming dataset:
```{r}
had <- c("AUS", "AUT", "BEL", "CAN", "CZE", "DNK", "FIN", "DEU", "GRC", "ISL", "IRL", "ITA", "JPN", "KOR", "MEX", "NZL", "NOR", "POL", "PRT", "SVK", "ESP", "SWE", "CHE", "TUR", "GBR", "CHL", "EST", "ISR", "SVN", "ZAF")
length_all_countries = length(had) + length(didnothave)
data <- data.frame(indicator = numeric(), gdp.per.capita = numeric(),
           gni.per.capita = numeric(), inflation = numeric(), spending = numeric(),  deficit = numeric(), unemloyment = numeric(),  stringsAsFactors = FALSE)
counter = 1
for(i in had){
  data[counter, 1] = 1
  data[counter, 2] = gdp[which(gdp$LOCATION==i & gdp$TIME == 2016)[2], 7]
  data[counter, 3] = gni[which(gni$LOCATION==i & gni$TIME == 2016)[2], 7]
  data[counter, 4] = inflation[which(inflation$LOCATION==i & inflation$TIME == 2016 & inflation$SUBJECT =="TOT")[1], 7]
  data[counter, 5] = spending[which(spending$LOCATION==i & spending$TIME == 2016 & spending$MEASURE == "MLN_USD"), 7]
  data[counter, 6] = deficit[which(deficit$LOCATION==i & deficit$TIME == 2016), 7]
  data[counter, 7] = unemloyment[which(unemloyment$LOCATION==i & unemloyment$TIME == 2016 & unemloyment$SUBJECT == "TOT"), 7]
  counter = counter + 1
}

```

```{r}
didnothave <- c("FRA", "COL", "RUS", "LVA", "LTU")
# CHN and IDN BRA CRI - no unemploy SAU -- no spending ARG BGR HRV CYP MLT ROU + + ZMBMKD(GNI INFl)--> no deficit 
for (i in didnothave) {
  data[counter, 1] = 0
  data[counter, 2] = gdp[which(gdp$LOCATION==i & gdp$TIME == 2016)[2], 7]
  data[counter, 3] = gni[which(gni$LOCATION==i & gni$TIME == 2016)[2], 7]
  data[counter, 4] = inflation[which(inflation$LOCATION==i & inflation$TIME == 2016 & inflation$SUBJECT =="TOT")[1], 7]
  data[counter, 5] = spending[which(spending$LOCATION==i & spending$TIME == 2016 & spending$MEASURE == "MLN_USD"), 7]
  data[counter, 6] = deficit[which(deficit$LOCATION==i & deficit$TIME == 2016), 7]
  data[counter, 7] = unemloyment[which(unemloyment$LOCATION==i & unemloyment$TIME == 2016 & unemloyment$SUBJECT == "TOT"), 7]
  counter = counter + 1
}
```
```{r}
data
```

```{r}
formula <- as.formula('indicator ~ gdp.per.capita + gni.per.capita + inflation + spending + deficit + unemloyment ')
#Logit
logit <- glm(formula, data , family = binomial('logit'))
summary(logit)
# df$fit <- predict(logit, df)
```
So we see that there are only 3 significant macroeeconomic factors that influence whether a country had a reform at 2016 or not.

To estimate the relative amount of information lost by a given model we will use AIC. 
```{r}
# install.packages("stats")
library("stats")
AIC(logit)
```

We see that AIC is relatively big. Let`s reduce the number of independent variables in order to get better results (as far as AIC is ...)
```{r}
data2 = data.frame(data[1], data[2], data[3], data[7])
data2
formula2 <- as.formula('indicator ~ gdp.per.capita + gni.per.capita + unemloyment ')
#Logit
logit2 <- glm(formula2, data2 , family = binomial('logit'))
summary(logit2)
logLik(logit2)
AIC(logit2)
```
We see that our model is badly explained by macroeconomic factors, that is why we will not use this one in order to make predictions about Ukraine. 
Let's enrich our sample with data points and a little bit different dataset (in order to enrich it with countries that did not have data about for example deficit and that is why were not included in the model)
```{r}
data2 <- data.frame(indicator = numeric(), gdp.per.capita = numeric(), spending = numeric(),  stringsAsFactors = FALSE)
counter = 1
for(i in had){
  data2[counter, 1] = 1
  data2[counter, 2] = gdp[which(gdp$LOCATION==i & gdp$TIME == 2016)[2], 7]
  data2[counter, 3] = spending[which(spending$LOCATION==i & spending$TIME == 2016 & spending$MEASURE == "MLN_USD"), 7]
  counter = counter + 1
}

didnothave <- c("FRA", "COL", "RUS", "LVA", "LTU", "ARG", "BGR", "HRV", "CYP", "MLT", "ROU", "ZMB","MKD")
# CHN and IDN BRA CRI - no unemploy SAU -- no spending ARG BGR HRV CYP MLT ROU + + ZMBMKD(GNI INFl)--> no deficit 

for (i in didnothave) {
  data2[counter, 1] = 0
  data2[counter, 2] = gdp[which(gdp$LOCATION==i & gdp$TIME == 2016)[2], 7]
  data2[counter, 3] = spending[which(spending$LOCATION==i & spending$TIME == 2016 & spending$MEASURE == "MLN_USD"), 7]
  counter = counter + 1
}
data2
formula2 <- as.formula('indicator ~ gdp.per.capita  + spending ')
#Logit
logit2 <- glm(formula2, data2, family = binomial('logit'))
summary(logit2)
logLik(logit2)
AIC(logit2)
```
Now as AIC has significantly decreased (due to increasing of likelihood) as well as results become more significant (possibly due to removing n.o. independent variables which were correlated as well adding more dataoiints about countries where there were no social government insurance).



